WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2014-06-02 ms.tex
Specify prior

A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components:
%
\begin{equation}
    \boldsymbol\theta = \begin{array}{l@{}l}
            \left[
            \begin{array}{c}
                \alpha \\
                \delta \\
                D_\mathrm{L} \\
                t_\oplus \\
                u \\
                \psi \\
                \phi_c \\
                \hdashline[1pt/1pt]
                m_1 \\
                m_2 \\
                \mathbf S_1 \\
                \mathbf S_2
            \end{array}
            \right] &
            \begin{array}{l}
                \left.
                \begin{array}{p{4cm}}
                    right ascension \\
                    declination \\
                    luminosity distance \\
                    arrival time at geocenter \\
                    $\cos(\textrm{inclination angle})$ \\
                    polarization angle \\
                    coalescence phase \\
                    \hdashline[1pt/1pt]
                \end{array}
                \quad \right\} \textnormal{extrinsic parameters}, \boldsymbol\theta_\mathrm{ex} \\
                \left.
                \begin{array}{p{4cm}}
                    first component's mass \\
                    second component's mass \\
                    first component's spin \\
                    second component's spin
                \end{array}
                \quad \right\} \textnormal{intrinsic parameters}, \boldsymbol\theta_\mathrm{in}
            \end{array}
        \end{array}
\end{equation}
%
This list of parameters involves some simplifying assumptions. Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should circularize due to tidal interaction~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz. Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\cite{PhysRevD.81.123016}. Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{BNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.

@@

We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$. The distance prior is given by a user-selected power of distance, ${D_\mathrm{L}}^m$, where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a scale-invariant prior that is uniform in $\log D_\mathrm{L}$. The priors are bounded as follows:
%
\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}%
%
\begin{eqnarray*}
	D_\mathrm{L} &\in& [D_{\mathrm{L},\mathrm{min}}, D_{\mathrm{L},\mathrm{max}}] \\
	t_\oplus &\in& [-\infty, \infty] \\
	u &\in& [-1, 1] \\
	\psi &\in& [0, \pi/2)
\end{eqnarray*}

>>>>

A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components:
%
\begin{equation}
    \boldsymbol\theta = \begin{array}{l@{}l}
            \left[
            \begin{array}{c}
                \alpha \\
                \delta \\
                D_\mathrm{L} \\
                t_\oplus \\
                \iota \\
                \psi \\
                \phi_c \\
                \hdashline[1pt/1pt]
                m_1 \\
                m_2 \\
                \mathbf S_1 \\
                \mathbf S_2
            \end{array}
            \right] &
            \begin{array}{l}
                \left.
                \begin{array}{p{4cm}}
                    right ascension \\
                    declination \\
                    luminosity distance \\
                    arrival time at geocenter \\
                    inclination angle \\
                    polarization angle \\
                    coalescence phase \\
                    \hdashline[1pt/1pt]
                \end{array}
                \quad \right\} \textnormal{extrinsic parameters}, \boldsymbol\theta_\mathrm{ex} \\
                \left.
                \begin{array}{p{4cm}}
                    first component's mass \\
                    second component's mass \\
                    first component's spin \\
                    second component's spin
                \end{array}
                \quad \right\} \textnormal{intrinsic parameters}, \boldsymbol\theta_\mathrm{in}
            \end{array}
        \end{array}
\end{equation}
%
This list of parameters involves some simplifying assumptions. Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should circularize due to tidal interaction~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz. Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\cite{PhysRevD.81.123016}. Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{BNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.

@@

We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}_\mathrm{in}$.

The arrival time prior is connected to the origin of the detector coordinate system. Given the Earth-fixed coordinates of the detectors $\mathbf{n}_i$ and the arrival times $\tau_i$, we compute their averages weighted by the timing uncertainty formula:
%
\begin{equation*}
    \langle \mathbf{n} \rangle = \frac{
        \displaystyle
        \sum_i \frac{\mathbf{n}_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
    }{
        \displaystyle
        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
    },
    \qquad
    \langle \hat\tau \rangle = \frac{
        \displaystyle
        \sum_i \frac{\hat\tau_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
    }{
        \displaystyle
        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
    }.
\end{equation*}
%
Then, we subtract off these means:
%
\begin{equation*}
    \mathbf{n}_i \leftarrow \mathbf{n}_i - \langle \mathbf{n} \rangle,
    \qquad
    \hat\tau_i \leftarrow \hat\tau_i - \langle \hat\tau \rangle.
\end{equation*}
%
In these coordinates relative to the weighted detector array barycenter, the arrival time prior is uniform in $-T \leq t \leq T$, with $T = \max\limits_i |\mathbf{n}_i| / c + 5\,\textrm{ms}$.

The distance prior is a user-selected power of distance,
%
\begin{equation*}
    p(D_\mathrm{L}) \propto \cases{
        {D_\mathrm{L}}^m & if $D_{\mathrm{L},\mathrm{min}} < D_\mathrm{L} < D_{\mathrm{L},\mathrm{max}}$ \\
        0 & otherwise
    }
\end{equation*}
%
where $m=2$ for a prior that is uniform in volume, and $m=-1$. If a distance prior is not specified, the default is uniform in volume out to the maximum SNR=4 horizon distance:
%
\begin{equation*}
    m = 2,
    \qquad
    D_{\mathrm{L},\mathrm{min}} = 0,
    \qquad
    D_{\mathrm{L},\mathrm{max}} = \frac{1}{4} \max_i D_{1,i}.
\end{equation*}

Finally, the prior is uniform in $-1 \leq \cos\iota \leq 1$ and $0 \leq \psi < 2\pi$.
